---
title: "2_Processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("variancePartition")

library(Seurat)
library(ggplot2)
library(dplyr)
library(ape)
library(cowplot)
library(Matrix)
library(variancePartition)
#BiocManager::install("MAST")
library(MAST)
#BiocManager::install('glmGamPoi')
library(glmGamPoi) #for sct
library(tibble)

#install.packages('devtools')
#devtools::install_github('immunogenomics/presto')
library(presto)
install.packages("R.utils")
#install.packages("remotes")
#remotes::install_github("satijalab/seurat-wrappers")
library(SeuratWrappers)

#install.packages("UpSetR")
library(UpSetR)
#install.packages("tidyverse")
library(tidyverse)
```

#Load filtered seurat object
```{r}
afterqc = "/Users/nathanewing-crystal/Ari Molofsky Lab Dropbox/Ari Molofsky Lab Team Folder/Nathan Ewing-Crystal/RNASeq/Lesion snRNASeq 3 Cxcl12/seurat/Data/Seurat/cxcl12_lesion_snRNASeq_filtered.RData"
load(afterqc)
sobject = sobject.merge
rm(sobject.merge)
```

```{r}
#Specify your preferred directory for all input + output
dir = "/Users/nathanewing-crystal/Ari Molofsky Lab Dropbox/Ari Molofsky Lab Team Folder/Nathan Ewing-Crystal/RNASeq/Lesion snRNASeq 3 Cxcl12/seurat"
#Specify exactly where your seurat files live
datafolder = "Data/Seurat"

project<-"cxcl12_lesion_snRNASeq"  

#set up folders
QCfolder = "QC"
Plotfolder = "Plots"

#metrics you want to look at for QC
m = c("nCount_RNA","nFeature_RNA","percent.mito","percent.ribo")

#You can add in housekeeping genes if you want them to be scaled always; otherwise set to "NULL"
add.genes = NULL

#Choose what to regress out - could be age, sex, or any metadata column
regress = c("nCount_RNA","percent.mito")

#Decide whether to use SCtransform (can take a long time with big datasets; generally recommended)
sct = TRUE

#How many genes do you want scaled to use for UMAP/clustering? 
ngenes = 6000

#Which principal components do you want to calculate on? This is a default setting, change if one of the pc's is for something you think is a technical error (i.e. HSP, RP, etc)
pcs = c(1:30)

#clustering resolution; the last number will be saved as "seurat_clusters" in metadata
res = c(1.5,1.0,0.75, 0.5, 0.25)

#Important genes to determine your cells of interest
igenes = c("Aldh1l1", "Gfap", "P2ry12", "Ptprc", "Itgam", "Col1a2", "Cthrc1", "Cd8a", "Cd4", "Rbfox3", "Epcam", "Itgam")

#metadata dimensions you want to cluster on
dims = c("seurat_clusters","sample", "timepoint","genotype")

#sexgenes to remove from dataset
sexgenes = c("Xist", "Tsix", "Ddx3y", "Eif2s3y", "Fkbp5", "Ddit4", "Uty")

#edit to include all desired celltypes to subset on later; if not subsetting, set to "all"
keep = "all"
#c("cd8"etc)

#make a unique name; maybe this is the celltype you've already subset, or the age you're looking at, etc.
iterationname = "scaled_processed"

#Establish cutoffs for heatmaps
pval = 1e-3 #max p-value for significance
lfc = 0.2 #minimum log fold change
minpct = 0.05 #if you want to subset by percent cells in that cluster expressing the gene
maxpct = 1
single = F #should each gene be a marker of a single cluster only
hgenes = 3 #how many genes should be in the heatmap per cluster
ncells = 100 #max # of cells per heatmap column
column = "seurat_clusters" #division you care about

#Variance partition: Remove genes only associated with a technical variable or sex of the mouse
variance = T
```

Function to print multiple graphs: 
```{r}
PrintSeuratGraph = function(namecard = "a",seurat_object = sobject,graphtype = "feature",feature = NULL,group = NULL,split=NULL,cellnames=NULL, label=FALSE){
  if (!is.null(cellnames)){
    #Idents(seurat_object) = cellnames[1]
    #cells = colnames(seurat_object)[Idents(seurat_object) %in% cellnames[1:length(cellnames)]]} 
    cells = cellnames}
  else {cells = cellnames}
  if (graphtype == "feature"){
    graph = FeaturePlot(seurat_object,features = feature,split.by = split, cells = cells,cols = c("lightyellow","darkred"))
  }
  if (graphtype == "violin"){
    graph = VlnPlot(seurat_object,features = feature, pt.size = 0.1, idents = cellnames[2:length(cellnames)],group.by = group, split.by = split)
  }
  if (graphtype == "dim"){
    graph = DimPlot(seurat_object,cells = cells, group.by = group, split.by = split, label = label)
  }
  name = paste0(feature,"_",graphtype,namecard,".eps")
  graph
  setEPS()
  postscript(file.path(dir,Plotfolder,name))
  print(graph)
  dev.off()
}
```

Find variable features, normalize, scale, run PCA, clustering, umap
The following is the standard method of normalization and scaling. Interchangeable with the next chunk. Run both, you should have already specified which to use with "sct = T or F". Will take 5-20 minutes to run. 
```{r,results = 'hide',eval = FALSE}
if (!sct){
  sobject <- NormalizeData(sobject,normalization.method = "LogNormalize", scale.factor = 10000)
  sobject<-FindVariableFeatures(sobject, selection.method = "vst", nfeatures = ngenes)
  
  all.genes<-rownames(sobject)
  var.genes = VariableFeatures(sobject)
  add.genes = add.genes[!add.genes %in% var.genes]
  
  any(add.genes %in% var.genes)
  scalegenes = c(var.genes,add.genes)
  VariableFeatures(sobject) = scalegenes
  sobject<-ScaleData(sobject,features = VariableFeatures(sobject), vars.to.regress = regress)
}
```

Alternative: SCTransform (great for smaller datasets)
```{r, results='hide',eval = FALSE}
if (sct){
  sobject <- SCTransform(sobject, vars.to.regress = regress, verbose = TRUE,variable.features.n = ngenes,conserve.memory = T)
}
```

Show most variable genes
```{r}
labels <- c(head(VariableFeatures(sobject),10),add.genes)
#Not working 2/2/25
plot1 = VariableFeaturePlot(sobject)
LabelPoints(plot=plot1, points = labels, repel = F, xnudge = 0.1, ynudge = 0.5)
```
Run "Variance Partition" - this will remove the genes most associated with sex (you can choose any column)

Takes a really long time if you load all genes
Individual and Tissue are both categorical,so model them as random effects; Note the syntax used to specify random effects

```{r,results= "hide",eval = FALSE}
if (variance){
  library(variancePartition)
  var.genes<-VariableFeatures(sobject)
  geneExpr = GetAssayData(sobject,slot = "data")
  geneExpr = as.matrix(geneExpr[rownames(geneExpr)%in% var.genes,])
  geneExpr = geneExpr[rowSums(geneExpr)>0,]
  info = sobject@meta.data
  info$nCount_RNA = info$nCount_RNA/sum(info$nCount_RNA)
  form <- ~ percent.mito + percent.ribo  + (1|sex) 
      # + (1|timepoint) + (1|condition) were taking too long
      # + nCount_RNA was making model fail and not included in later removals, so excluded
  varPart <- fitExtractVarPartModel(geneExpr, form, info )
  varPartSave = as.data.frame(varPart)
  write.csv(varPartSave,file = file.path(dir,"Spreadsheets",paste0(Project(sobject),"_variancepartition.csv")))
}
```

```{r}
varPart = read.csv(file.path(dir,"Spreadsheets",paste0(Project(sobject),"_variancepartition.csv")),stringsAsFactors = F,row.names = 1)
```

```{r}
if (variance){
  library(variancePartition)
  # sort variables (i.e. columns) by median fraction of variance explained
  vp <- sortCols(varPart )
  
  #order on each column
  vs = varPart[order(varPart$sex,decreasing = T),]
  vm = varPart[order(varPart$percent.mito,decreasing = T),]
  vr = varPart[order(varPart$percent.ribo,decreasing = T),]
  #vc = varPart[order(varPart$nCount_RNA,decreasing = T),]
  
  # Bar plot of variance fractions for the first 10 genes
  print(plotVarPart(varPart))
  print(plotPercentBars( vm[1:50,] ))
  print(plotPercentBars( vs[1:50,] ))
  plotPercentBars( vr[1:50,] )
  #print(plotPercentBars( vc[1:50,] ))
  
        
  # violin plot of contribution of each variable to total variance
  setEPS()
  postscript(file.path(dir,"QC",paste0(project,"vln_variance.eps")))
  print(plotVarPart(varPart))
  dev.off()
  
  setEPS()
  postscript(file.path(dir,"QC",paste0(project,"sex_variance.eps")))
  print(plotPercentBars( vs[1:50,] ))
  dev.off()
  
  setEPS()
  postscript(file.path(dir,"QC",paste0(project,"mito_variance.eps")))
  print(plotPercentBars( vm[1:50,] ))
  dev.off()
  
  setEPS()
  postscript(file.path(dir,"QC",paste0(project,"ribo_variance.eps")))
  print(plotPercentBars( vr[1:50,] ))
  dev.off()
  
 #  setEPS()
 #  postscript(file.path(dir,"QC",paste0(project,"ncount_variance.eps")))
 #  print(plotPercentBars( vc[1:50,] ))
 #  dev.off()
}
```

```{r}
if (variance){
  sexgenes = rownames(vs)[1:5]
  mitogenes = rownames(vm)[1:10]
  remove.genes = c(sexgenes,mitogenes)
  var.genes = VariableFeatures(sobject)
  var.genes = var.genes[!var.genes %in% remove.genes]
  VariableFeatures(sobject) = var.genes
}
```


Run PCA analysis and show elbow plot
```{r}
sobject <- RunPCA(sobject,features = VariableFeatures(sobject),npcs = 50, verbose = FALSE)
ElbowPlot(sobject,ndims = 50, reduction = "pca")
print(sobject[["pca"]], dims = 1:20, nfeatures = 5)
```

Once you are satisfied with pc's, run clustering: 
```{r}
sobject<-RunUMAP(sobject,reduction = "pca",dims = pcs, verbose = F)
sobject<-FindNeighbors(sobject,dims=pcs,verbose=F)
sobject<-FindClusters(sobject,verbose=F,resolution = res)
sobject$seurat_clusters = sobject$SCT_snn_res.0.25
```

Plot important objects; check parameters before moving forward, evaluate QC, clustering
```{r}
for (dim in dims){
  print(DimPlot(sobject,group.by = dim, label = T))
}

DimPlot(sobject, split.by = "sample")
FeaturePlot(sobject,igenes, order=T, slot="counts")
FeaturePlot(sobject,m)

#Build a clustering tree
Idents(sobject) = "seurat_clusters"
sobject= BuildClusterTree(sobject,dims = pcs)
tree = sobject@tools$BuildClusterTree
plot.phylo(tree, use.edge.length = T, direction = "rightwards")
```

*Note that the chunk below requires you to manually decide which clusters to keep or exclude. The violin plots will help decide based on marker genes. You can also choose to annotate (as many types as you want) and skip subsetting.*
```{r}
#Edit this part carefully. You can add any number of types. Each cluster can only be one type. 

type0 = c(-1)
name0 = "Fibroblast_1"
type1 = c(0)
name1 = "Myeloid_1"
type2 = c(1)
name2 = "Fibroblast_2"
type3 = c(2)
name3 = "Neuron_1"
type4 = c(3)
name4 = "Neuron_2"
type5 = c(4)
name5 = "Astrocyte_1"
type6 = c(5)
name6 = "Neuron_3"
type7 = c(6)
name7 = "Neuron_4"
type8 = c(7)
name8 = "Neuron_5"
type9 = c(8)
name9 = "Neuron_6"
type10 = c(9)
name10 = "Neuron_7"
type11 = c(10)
name11 = "Oligo"
type12 = c(11)
name12 = "Fibroblast_3"
type13 = c(12)
name13 = "Neuron_8"
type14 = c(13)
name14 = "Myeloid_2"
type15 = c(14)
name15 = "Neuron_9"
type16 = c(15)
name16 = "Endothelial"
type17 = c(16)
name17 = "T_cells"
type18 = c(17)
name18 = "Neuron_9"
type19 = c(18)
name19 = "Neuron_10"
type20 = c(19)
name20 = "Astrocyte_2"
type21 = c(20)
name21 = "Unassigned"

#Initialize the cluster levels as a vector and replace the cluster levels with the appropriate name. 
clusters = as.factor(sobject$seurat_clusters)
type = levels(clusters)
type[type0+2] = name0
type[type1+2] = name1
type[type2+2] = name2
type[type3+2] = name3
type[type4+2] = name4
type[type5+2] = name5
type[type6+2] = name6
type[type7+2] = name7
type[type8+2] = name8
type[type9+2] = name9
type[type10+2] = name10
type[type11+2] = name11
type[type12+2] = name12
type[type13+2] = name13
type[type14+2] = name14
type[type15+2] = name15
type[type16+2] = name16
type[type17+2] = name17
type[type18+2] = name18
type[type19+2] = name19
type[type20+2] = name20
type[type21+2] = name21

levels(clusters) = type

#Add a metadata column
sobject$celltype = clusters

#check the celltype assignment for accuracy
table(sobject$celltype,sobject$seurat_clusters)

#Check them against your marker genes
VlnPlot(sobject,igenes,group.by = "celltype",pt.size = 0.01)

#add a metadata column labelling each cluster
sobject$celltypecluster = paste0(sobject$celltype,"-",sobject$seurat_clusters)
```

Save the tree
```{r}
Idents(sobject) = column
sobject= BuildClusterTree(sobject,dims = pcs)
tree = sobject@tools$BuildClusterTree
setEPS()
postscript(file.path(dir,"Trees",paste0(project,"_",iterationname,"_tree.eps")))
plot.phylo(tree, use.edge.length = T, direction = "rightwards")
dev.off()
```

#After clustering, analysis should be performed on normalized/scaled "RNA" slot
```{r}
DefaultAssay(sobject) <- "RNA"
sobject = JoinLayers(sobject) # new in this analysis - collapse datasets to create counts and data layers (seurat v5) 
sobject <- NormalizeData(sobject)
sobject <- ScaleData(sobject)
```
Save the clustered dataset. Overwrite the existing, subset dataset. 
```{r,eval = FALSE}
save(sobject,file = file.path(dir,datafolder,paste0(project,"_",iterationname,".RData")))
```

Block to print multiple graphs: 
```{r}
Plotfolder = "Plots"
name = paste0(project,"_",iterationname)
genes = igenes
features = m
groups = c(dims, "celltype","celltypecluster", "genotype", "timepoint", "sex", "microanatomy")
genes = genes[genes %in% rownames(sobject)]

for(feature in genes){
  PrintSeuratGraph(namecard = name,graphtype = "feature",feature = feature)
}

for(feature in features){
  PrintSeuratGraph(namecard = name,graphtype = "feature",feature = feature)
}

#dim plots for clustering
for(group in groups){
  PrintSeuratGraph(namecard = name,graphtype = "dim",group = group, feature = group)
  PrintSeuratGraph(namecard = paste0(name,"_labeled_"),graphtype = "dim",group = group, feature = group, label = TRUE)
  PrintSeuratGraph(namecard = paste0(name, "_split"),graphtype = "dim",feature = group, split = group)
}

#celltype without legend
graph = DimPlot(sobject, group.by = "celltype", label=TRUE) + ggplot2::theme(legend.position = "none")
setEPS()
postscript(file.path(dir,Plotfolder,"dimplot_nolegend.eps"))
print(graph)
dev.off()

PrintSeuratGraph(namecard = paste0(name, "_celltype-genotype"), graphtype = "dim", group = "celltype", split = "genotype")

#violin plots
for(feature in c(genes,features)){
  PrintSeuratGraph(namecard = name,graphtype = "violin",feature = feature,group = "seurat_clusters")
}
```

Subset the data if necessary; will only run if "keep" includes some celltype

Subset the data to include only your cells of interest: 
```{r}
iterationname = "subset"

#create immune, fibroblast, and astrocyte subset Seurat objects for reclustering
celltype = levels(as.factor(sobject$celltype))
myeloid_keep = celltype[c(2,15)]
fibroblast_keep = celltype[c(1,3,13)]
t_keep = celltype[c(18)]
neuron_keep = celltype[c(4,5,7,8,9,10,11,14,16,19)]
astro_keep = celltype[c(6,20)]

sobject.sub.myeloid = subset(sobject,subset = celltype %in% myeloid_keep)
save(sobject.sub.myeloid,file = file.path(dir,datafolder,paste0(project,"_",iterationname,"_myeloid.RData")))
sobject.sub.fib = subset(sobject,subset = celltype %in% fibroblast_keep)
save(sobject.sub.fib,file = file.path(dir,datafolder,paste0(project,"_",iterationname,"_fib.RData")))
sobject.sub.t = subset(sobject, subset = celltype %in% t_keep)
save(sobject.sub.t,file = file.path(dir,datafolder,paste0(project,"_",iterationname,"_t.RData")))
sobject.sub.neuron = subset(sobject, subset = celltype %in% neuron_keep)
save(sobject.sub.neuron,file = file.path(dir,datafolder,paste0(project,"_",iterationname,"_neuron.RData")))
sobject.sub.astro = subset(sobject, subset = celltype %in% astro_keep)
save(sobject.sub.astro,file = file.path(dir,datafolder,paste0(project,"_",iterationname,"_astro.RData")))
```

